Using species distribution models to direct native seed collection efforts

### 2024-09-13 scripts/Results.Rmd
conceptual figs

@@

ggplot(suitability_amounts, aes(x = Percent, y = Records, color = Species, group = Species)) +
  geom_line(lty = 1) +
  scale_color_grey(guide = 'none') + 
  geom_ribbon(data = suitability_means, aes(ymin = lower, ymax = upper, x = Percent),
              fill = '#73BA9B', inherit.aes = F, alpha = 0.5) + 
  geom_line(data = suitability_means, 
            aes(x = Percent, y =  mean, color = 'a'), color = '#093824', inherit.aes = F, lwd = 1) + 
  geom_line(data = fp, aes(x = suitability*100, y = obs), color = '#FF2C55', inherit.aes = F,  lwd = 1)

@@

rm(linecols)

hypothetical_curves
```

>>>>

habitat_specialist <- data.frame(
  suitability = c(
    seq(from= 0, to  = 0.1, length.out = 70), 
    seq(from= 0.1, to  = 0.4, length.out = 15), 
    seq(from= 0.4, to  = 0.8, length.out = 10),
    seq(from= 0.8, to  = 1, length.out = 5)
  ),
  percent = 1:100
)

habitat_generalist <- data.frame(
  suitability = c(
    seq(from= 0, to  = 0.5, length.out = 40), 
    seq(from= 0.6, to  = 0.7, length.out = 30), 
    seq(from= 0.7, to  = 0.9, length.out = 20),
    seq(from= 0.9, to  = 1, length.out = 10)
  ),
  percent = 1:100
)

ggplot() + 
  geom_smooth(data = habitat_specialist, aes(x = percent, y = suitability)) + 
  geom_smooth(data = habitat_generalist, aes(x = percent, y = suitability)) +
  theme(
    plot.title = element_text(hjust = 0.5),
    aspect.ratio = 1/1)

@@

suitability_means <- suitability_means |>
  mutate(Null_mod = mean*Percent)

ggplot(suitability_amounts,
       aes(x = Percent, y = Records, color = Species, group = Species)) +
  geom_line(lty = 1) +
  scale_color_grey(guide = 'none') + 
  geom_ribbon(data = suitability_means, aes(ymin = lower, ymax = upper, x = Percent),
              fill = '#73BA9B', inherit.aes = F, alpha = 0.5) + 
  geom_line(data = suitability_means, 
            aes(x = Percent, y =  mean, color = 'a'), 
            color = '#093824', inherit.aes = F, lwd = 1) + 
  geom_line(data = suitability_means, aes(x = Percent, y = Null_mod/100),
            color = 'black', inherit.aes = F,  lwd = 1) +
  geom_line(data = fp, aes(x = suitability*100, y = obs),
            color = '#FF2C55', inherit.aes = F,  lwd = 1) +

@@

rm(linecols)
```

```{r}
suitability_means[ which.min(abs(suitability_means$Percent-50)), 'mean'] # on average 50% of cells have lower probabilities of being suitable than 0.23

suitability_means[ which.min(abs(suitability_means$Percent-80)), 'mean'] # on average 80% of cells have lower probabilities of being suitable than 0.614

fp[ which.min(abs(fp$obs-0.2)), ] # 80% of all populations are found in areas with suitability > 65.2%

fp[ which.min(abs(fp$obs-0.5)), ] # 50% of all populations are found in areas with suitability > 83.6%

fp[ which.min(abs(fp$obs-0.8)), ] # the final 20 % of populations are found in areas with suitability > 94.6%
```

### 2024-09-13 scripts/Using species distribution models to direct native seed collection efforts.Rmd
conceptual figs

@@

@@

*Where were the populations crews ended up finding?* weird elbow plot)
How did presences and absences interact? (logistic regression)

```{r}
rm(recs_removed)
rm(list = ls()[ sapply( ls(), exists, mode="function")])

>>>>

## Ground Verification 
13 native seed collection crews had three products shared with them for planning their field work to scout populations for collections. 
1) All input training and test occurrence points used in generating the models.
2) All SDM models, in raster format, clipped to > 80% probability of suitable habitat.
3) All drainage basins ranked via an algorithm similar to the above, as vector data, but where the dis-contiguous patches were assigned ranks of '' and '' respectively. 
Some crews, notable those with less amounts of standardized monitoring projects in their areas, which adversely affect product 3, were also given raw SDM output rasters, i.e. the rasters not masked to <80% suitability.

Four crews, all located on the northern Colorado Plateau, were given a checklist for taxon specific presence-absence monitoring to begin in late August until the end of their terms in October.  
These 27 species were selected as they already had multiple presence records (>3), crews were asked to fill out these forms 8 times or more each, across their field offices as time permitted. 
This subset of taxa, and number of sampling events, were chosen so that each taxon could become an effect in a logistic regression model, and so that crews could collect these data relatively quickly.

@@

```{r Populations found across all crews}

```{r Presences and Absences from Colorado Plateau Crews}

```{r Populations relative to drainages}

\begin{wrapfigure}{R}{0.5\textwidth}  
 \begin{center}
    \includegraphics[width=.5\textwidth]{../plots/VariableImportance.png}  
  \caption{bird is the word} 
\end{center}
\end{wrapfigure}

@@

\begin{wrapfigure}{R}{0.35\textwidth}  
 \begin{center}
    \includegraphics[width=.35\textwidth]{../plots/Eval_metrics.png}  
  \caption{word to all my birds} 
\end{center}
\end{wrapfigure}

*Where were the populations crews ended up finding?* weird elbow plot)

How did presences and absences interact? (logistic regression)

Where were newly found populations in terms of drainage connections?

```{r}
rm(recs_removed)
rm(list = ls()[ sapply( ls(), exists, mode="function")])
```
